Index measures for oak decline severity using phenotypic descriptors

### 2020-02-10 manuscript/manuscript.Rmd
added section references

## Abstract {-}

@@

## Introduction

@@

## Materials and methods

### Phenotypic data collection

@@

### Phenotypic descriptor preparation

@@

### Composite descriptor calculation

@@

### Random forest analysis of inter-site discrimination

@@

### Tree size descriptor adjustment for age and inter site variance

@@

### Calculation of indexes for oak decline severity

@@

### Descriptor contribution to decline indexes

To investigate the descriptor contributions to the decline indexes, supervised random forest regression models were trained for each decline index using the phenotypic descriptors for all 174 trees.
These models will subsequently be referred to as $M_{PDI}$ and $M_{DAI}$ for the PDI and DAI respectively. 
The optimal number of trees ($ntree$) and number randomly of variables ($mtry$) parameters for random forest model were tuned prior to training using the values of 10, 100, 1,000, 10,000 for ntree and a range of 2-26 at intervals of 4 for the mtry parameter. 
The combination parameters giving the lowest percent mean absolute error (PMAE) were selected and set as follows: for the PDI ($M_{PDI}$): `r glue('$ntree$ = {prettyNum(PDI_rf_tune_params$ntree,big.mark = ',')}, $mtry$ = {PDI_rf_tune_params$mtry}, PMAE = {PDI_rf_tune_params$PMAE %>% signif(digits = 3)}%')`; for the DAI ($M_{DAI}$): `r glue('$ntree$ = {prettyNum(DAI_rf_tune_params$ntree,big.mark = ',')}, $mtry$ = {DAI_rf_tune_params$mtry}, PMAE = {DAI_rf_tune_params$PMAE %>% signif(digits = 3)}%')` (Supplementary figure 2).
Overall descriptor importance contributions to each of the indexes using the mean decrease in accuracy measure as described in Section #.

### Local descriptor contribution to index models

@@

### Index model simulations to investigate interactions of key descriptors

@@

## Results

### Phenotypic differences between site locations and trees size descriptor adjustments

A preliminary analysis using supervised random forest classification could separate trees based on their location (Supplementary  figure 3a) with an average *margin* value of `r site_rf_margin$Margin %>% signif(digits = 3)`.
The descriptors that contributed most to this discrimination were those that were related to trees size such as DBH, total height and crown radius and likely reflect the varied tree age and planting density across the nine sites (Supplementary figure 3b). 
To counter this, effect sizes were adjusted based on the site means for each tree size descriptor to remove inter-site variability but maintain intra-site variability as described in Section #.
This ensures that the subsequent comparisons of these descriptors in the context of oak decline are not confounded by these identified site differences.
After the application of the tree size adjustment, the average random forest margin value was substantially reduced to `r site_rf_post_correction_margin$Margin %>% signif(digits = 3)`.
This negative value indicates that the adjustment was able to remove the majority of the between site variance and that the random forest models were on average unable to correctly classify the site locations of the trees.

### Oak decline severity indexes of manually assigned decline statuses

@@

### Key descriptors contributing to the decline indexes

@@

### Local descriptor contributions and trends within the PDI and DAI models

@@

### Key descriptor interactions within the decline index models

The complex descriptor interactions within the index models ($M_{PDI}$, $M_{DAI}$) were investigated by simulating phenotypic data and generating the index values from the supervised index random forest regression models. 
The index values generated by the models from these simulated phenotypic data were then visualised as model response surface plots, from which descriptor interactions within the indexes can be interpreted. 
Figures 3 and 4 show the simulated index value response surfaces under a variety of different circumstances for the key phenotypic descriptors described in Section ## that contribute highly to the PDI and DAI.

@@

## Discussion

### Suitability of the decline indexes for describing decline severity and type

@@

### The machine learning approach

@@

### Applications

@@

## Conclusions

@@

## Acknowledgements

@@

## References

>>>>

# Abstract {-}

@@

# Introduction {#introduction}

@@

# Materials and methods {#materials-and-methods}

## Phenotypic data collection {#data-collection}

@@

## Phenotypic descriptor preparation {#method-decriptor-preparation}

@@

## Composite descriptor calculation {#method-composite-descriptors}

@@

## Random forest analysis of inter-site discrimination {#method-inter-site-differences}

@@

## Tree size descriptor adjustment for age and inter site variance {#method-adjustment}

@@

## Calculation of indexes for oak decline severity {#method-index-calculation}

@@

## Descriptor contribution to decline indexes {#method-descriptor-contribution}

To investigate the descriptor contributions to the decline indexes, supervised random forest regression models were trained for each decline index using the phenotypic descriptors for all 174 trees.
These models will subsequently be referred to as $M_{PDI}$ and $M_{DAI}$ for the PDI and DAI respectively. 
The optimal number of trees ($ntree$) and number randomly of variables ($mtry$) parameters for random forest model were tuned prior to training using the values of 10, 100, 1,000, 10,000 for ntree and a range of 2-26 at intervals of 4 for the mtry parameter. 
The combination parameters giving the lowest percent mean absolute error (PMAE) were selected and set as follows: for the PDI ($M_{PDI}$): `r glue('$ntree$ = {prettyNum(PDI_rf_tune_params$ntree,big.mark = ',')}, $mtry$ = {PDI_rf_tune_params$mtry}, PMAE = {PDI_rf_tune_params$PMAE %>% signif(digits = 3)}%')`; for the DAI ($M_{DAI}$): `r glue('$ntree$ = {prettyNum(DAI_rf_tune_params$ntree,big.mark = ',')}, $mtry$ = {DAI_rf_tune_params$mtry}, PMAE = {DAI_rf_tune_params$PMAE %>% signif(digits = 3)}%')` (Supplementary figure 2).
Overall descriptor importance contributions to each of the indexes using the mean decrease in accuracy measure as described in Section \@ref(method-inter-site-differences).

## Local descriptor contribution to index models

@@

## Index model simulations to investigate interactions of key descriptors

@@

# Results

## Phenotypic differences between site locations and trees size descriptor adjustments

A preliminary analysis using supervised random forest classification could separate trees based on their location (Supplementary  figure 3a) with an average *margin* value of `r site_rf_margin$Margin %>% signif(digits = 3)`.
The descriptors that contributed most to this discrimination were those that were related to trees size such as DBH, total height and crown radius and likely reflect the varied tree age and planting density across the nine sites (Supplementary figure 3b). 
To counter this, effect sizes were adjusted based on the site means for each tree size descriptor to remove inter-site variability but maintain intra-site variability as described in Section \@ref(method-adjustment).
This ensures that the subsequent comparisons of these descriptors in the context of oak decline are not confounded by these identified site differences.
After the application of the tree size adjustment, the average random forest margin value was substantially reduced to `r site_rf_post_correction_margin$Margin %>% signif(digits = 3)`.
This negative value indicates that the adjustment was able to remove the majority of the between site variance and that the random forest models were on average unable to correctly classify the site locations of the trees.

## Oak decline severity indexes of manually assigned decline statuses

@@

## Key descriptors contributing to the decline indexes {#results-descriptor-contirbutions}

@@

## Local descriptor contributions and trends within the PDI and DAI models

@@

## Key descriptor interactions within the decline index models

The complex descriptor interactions within the index models ($M_{PDI}$, $M_{DAI}$) were investigated by simulating phenotypic data and generating the index values from the supervised index random forest regression models. 
The index values generated by the models from these simulated phenotypic data were then visualised as model response surface plots, from which descriptor interactions within the indexes can be interpreted. 
Figures 3 and 4 show the simulated index value response surfaces under a variety of different circumstances for the key phenotypic descriptors described in Section \@ref(results-descriptor-contirbutions) that contribute highly to the PDI and DAI.

@@

# Discussion

## Suitability of the decline indexes for describing decline severity and type

@@

## The machine learning approach

@@

## Applications

@@

# Conclusions

@@

# Acknowledgements {-}

@@

# References
